Service Oriented ARC Storage; \\An Overview and Design

### 2008-12-04 doc/storage_system/ARCStorage_src/ARCStorage.tex
Implementing Alex' comments up to Testing and Discussion

\begin{document}
\pagestyle{headings} 
%% \title{Service Oriented ARC Storage; \\An Overview and Design}
\title{Service Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$  \\ $^{*}$Scientific
  Computing Division, IT Department, Uppsala University, Sweden
  \\ $^{\dagger}$Center for Information Technology, University of Oslo, Norway
  \\ $^{\ddagger}$Department of Physics,  University of Oslo, Norway
  \\  $^{\mathsection}$Department of Physics and Astronomy, Uppsala University,
  Sweden \\ $^{\mathparagraph}$National Information Infrastructure Development
  Institute, Hungary\\ \\  j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

\maketitle             
%% \begin{abstract}
%% The KnowARC project will deliver the next generation Advanced Resource
%% Connector Fall  2009. A major component of this
%% service oriented Grid middleware will be its storage system, the
%% ARC Storage, of which a
%% prototype is due Spring 2009. We will in this paper present an
%% overview of the storage system, its design and architecture. We will
%% also present some first, proof-of-concept test results, deploying the
%% the storage with a geographically distribution among three different countries.
%% ~\\~\\~
%% \end{abstract}
\begin{abstract}
There is an ever increasing need to utilize geographically distributed
hardware resources, both in terms of CPU and in terms of storage. The
service oriented architecture provides a natural framework for
managing these resources. The next generation Advanced Resource
Connector (ARC) is a service oriented Grid solution that will provide
the middleware to represent distributed resources in one simple
framework. In this paper, we will present an overview of the ARC
storage system, itself a set of services providing a self-healing,
flexible Grid storage solution. We will also present some first
proof-of-concept test results, deploying the storage system
distributed between three different countries. 
~\\~\\~
\end{abstract}
%-----------------------------------------------------------
\section{Introduction}
\label{Introduction}

The challenge of building a reliable, self-healing, fault-tolerant,
consistent data management system in a web scale is an interesting
task. Making the system work in a heterogeneous, distributed
environment like the Grid is even more interesting. An increasing
amount of applications demands not only increased CPU power, but also
vast amounts of storage space. The required storage space is not only
restricted to the duration in which the application runs; the data
should often be available for years afterwards, in certain cases even
for decades. Nowadays, we can easily find single Grid jobs, producing
gigabytes or even terabytes of data, ramping up the requirements of
storage systems to the petabyte scale and beyond. Taking into account
that hardware is inherently unstable\footnote{E.g., if one hard drive
  has an uptime of 10,000 hours, one of 1,000 hard drives will fail
  after 10 hours.}, the need for replicating, self-healing storage is evident. To
make the storage system useable to Grid users, the system must provide
reliable and secure file transfer protocols, a cataloging system and
secure storage of the data. Several projects and designs have emerged
to address such challenges \cite{Hoschek00datamanagement,DengWang}.

In the advent of the next generation Advanced Resource Connector (ARC)
Grid middleware \cite{NorduGridsite} (due during Fall 2009), we present 
the ARC storage system \cite{ARCStoragedesigndoc}. This distributed
storage system is designed to provide an easy
to use, flexible and scalable system that can offer native storage and
at the same time provide access to third-party solutions like dCache by
using the same, uniform interface. Being part of ARC, the
storage system is based on a service
oriented architecture, in which each major component of the
system runs as a separate service within the ARC Hosting Environment
Daemon (HED) \cite{HEDdesigndoc}. The HED service container gives the
capability of flexible replacement of the components as well as the
possibility to introduce modifications in the future. %% The HED service container gives
%% the capability of flexible replacement of the components as well as
%% the future modification/enhancement of the system.

This paper is organized as follows. After describing
related work in Section \ref{Related Work}, we give the bird's eye
view of the next generation Advanced Resource Connector (ARC) in
Section \ref{The Advance Resource Connector}. An overview of the
ARC Storage is given in Section \ref{The ARC Storage System}, while
the architecture of the storage system is elaborated in Section
\ref{Architecture of the ARC Storage}. In Section \ref{Testing and
  Discussion} we give some early, proof-of-concept results, before 
concluding in Section \ref{Conclusion and Future Work}.

@@

\textbf{dCache:} dCache is a storage system \cite{dCache} \cite{DSSWithdCache} which combines heterogeneous storage

@@

a joint effort between Desy  \cite{dCachesite}, Fermi National Laboratory \cite{FermiLab}, the  Nordic Data

@@

it has a relatively high threshold for understanding and
deployment. The ARC Storage, being a light-weight and flexible

@@

\textbf{BigTable:} Bigtable is a distributed storage system managing structured data in

@@

Google, in projects like web indexing, Google Earth and Google

@@

system. It allows  uniform access for connecting to heterogeneous

@@

Originally developed for use with the physics analysis tool root,

@@

high bandwidth environments. In contrast, the ARC Storage is optimized
for high latency, more suitable for the Grid environment.

@@

The next generation Advanced Resource Connector (ARC) Grid middleware \cite{NorduGridsite} \cite{KnowARCsite} consists of a set of pluggable components. These components are the
fundamental building blocks of the ARC services and clients. ARC
services are running inside a container called the Hosting Environment
Daemon (HED) and there are four kinds of pluggable components having well
defined tasks: Data Management Components are used to transfer the
data using various protocols, Message Chain Components are
responsible for the communication within clients and services as well as
between the clients and the services, ARC Client Components are
plug-ins used by the clients to connect to different Grid flavors and
Policy Decision Components are responsible for the security model
within the system.

To deliver the non-trivial quality of services required by the Grid,
there are a number of services running inside the HED. For example, Grid
job execution and management is handled by the A-REX service \cite{AREXdesigndoc}, policy
decision is the task for the Charon service, the ISIS service is
responsible for the information indexing, batch jobs submission is
handled by the Sched service etc. In the later sections our discussion
will focus on the architecture and the design of another major set
of services, i.e., the ARC storage.

@@

The ARC Storage consists of a set of SOAP based services residing
within HED. Together the services provide a self-healing, reliable, robust,
scalable, resilient and consistent data storage system. Data is
managed in a hierarchical global namespace with files and
subcollections grouped into collections\footnote{A concept very
  similar to files and directories in most common file systems.}.  A
dedicated root collection serves as a reference point for accessing the namespace. The hierarchy can then be
referenced using Logical Names. The global namespace is accessed
in the same manner as in local filesystems.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.8\columnwidth]{arc1-storage-services-with-gateway.pdf}
  \end{center}
  \caption{Schematic of the ARC Storage architecture. The figure shows
    the main services of ARC Storage and how they communicate. See
    text for further details.}
  \label{fig:arc1_storage_service}
\end{figure}

Being a service oriented architecture based system, the ARC
Storage consists of a set of services as shown in figure
\ref{fig:arc1_storage_service}. The services are as follows: The Bartender (B),
providing the high-level interface to the user; the Librarian (L),
handling the entire storage namespace, using the A-Hash (A-H) as a metadatabase; the
Gateway (G) providing means to access third-party storage systems; and
the Shepherd (S), the frontend for the physical storage node. 
See Section \ref{Architecture of the ARC Storage} for a detailed
discussion of the different services.   
The services communicates with each other through the
Message Chain Components in HED. The communication is
depicted by straight lines in figure \ref{fig:arc1_storage_service}.

The system supports file transfer through several transfer protocols,
with client side tools hiding technical
details, such as specifying protocols, port numbers
and so forth. To enable for fault-tolerance, the system implements
automatic file replication, where
the replicas of a file\footnote{In this paper, a file
denotes an entry in the global namespace, while replica denotes the
physical file stored on a storage node.} are always stored on separate
storage nodes\footnote{A
  storage node is a server with a Shepherd, a Shepherd backend
and some storage service.}. To ensure a resilient, self-healing system, the Shepherd
regularly sends heartbeats to one of the Librarians. If the Shepherd fails
to send a heartbeat one of the Librarians will automatically initiate
re-replication of the replicas between the other Shepherds.

In the default implementation the services will provide a
full-featured and consistent data storage system using files as an
atomic unit. The scope of the ARC storage system in the foreseen
future is restricted to provide support for file-based data services.
%% there are currently no plans for intrinsic support for database
%% access.

@@

External grid middleware components can
access the ARC storage system by adapting and using ARC data service
interfaces directly. ARC will provide interface
components that communicate via standard protocols like SRM, which
will give a single access point to the whole system.

@@

In a service oriented architecture
the role of each service is well defined. Available
objects\footnote{Files, collections, mount points, etc.} in the
system are identified by unique global IDs. These IDs are
categorized in three different categories according to the object type.

\begin{itemize} 
\item Each  file and collection have a unique ID called GUID.
\item Services are uniquely identified by using a serviceID. 
\item The Shepherds in the system identify their files using a referenceID.
\end{itemize}

Each object in the ARC Storage has a globally
unique ID. A collection contains files and other collections,
and each of these entries has a name unique within the collection very
much like entries in a standard directory on a local filesystem. Besides
files and collections the ARC Storage has a third type of entries
called mount-points which are references to the third-party storages
within the global namespace.

Replicas in a distributed storage system can have different states;
they can be broken, deleted, partially uploaded, etc. In the
ARC Storage, all replicas has assigned a state, some of which are
`\textbf{alive}' (if the replica passed the checksum test, and the
Shepherd reports that the storage node is healthy), `\textbf{invalid}'
(if the replica has wrong checksum, or the Shepherd claims it has no
such replica), `\textbf{creating}' (if the replica is in
the state of uploading).

The details of the
architecture are divided into two parts: First we will discuss the
details of the core components, second we will discuss some of the
important features provided by the system.

\subsection{Core Components} 
\subsubsection{Bartender}
\label{Bartender}
The Bartender service provides a high-level interface for the storage
system. Clients connect to the Bartender to create and remove
files, collections and mount-points using their Logical Names. The
Bartender communicates 
with the Librarian and Shepherd services to accomplish the clients
requests. However, the actual file data does not go through the
Bartender, instead file
transfers are directly performed between the storage nodes and the
clients. There could be any number of independent Bartender services
running in the system, providing high-availability and
load-balancing. 
\subsubsection{Librarian}
\label{librarian}
The Librarian manages the hierarchy and metadata of
files, collections and mount points, as well as the health information of the Shepherd
services. In addition, the Librarian handles
the information about registered Shepherd services. The Librarian receives
heartbeat messages from the Shepherds and change replica states automatically
if needed. The Librarian uses the A-Hash for consistent storing. This makes the
Librarian a stateless service, thus enabling the system to have any
number of independent
Librarian services, again providing high-availability in the
system.

@@

The Shepherd services run as front-ends on storage nodes. A Shepherd
service reports to a Librarian about the 
node's health state in terms of replicas. While the Bartender initiates
file transfers, the actual transfers goes directly between the Shepherd
and the clients.

@@

A-Hash is a hash table for consistently storing data in property-value
pairs. All metadata about files, collections, mount point, Shepherd's
health status, and so forth, is stored in the A-Hash. As the A-Hash is
the service storing the entire state of the storage system, its
absolutely crucial for the ARC Storage that the A-Hash is
consistent. The distribution and replication of this service is
therefore both challenging and needed.

@@

\subsection{Features}
\subsubsection{Heartbeat monitoring}
\label{Heartbeat and Replication}
In the proposed architecture, each Shepherd periodically send
heartbeats to a Librarian with information about replicas
whose state changed since the last heartbeat, and the corresponding
GUIDs. These heartbeats are then stored in the A-Hash, making them
visible to all the Librarians in the system. If any of the Librarians
notices that a Shepherd is late with its heartbeat, it will invalidate
all the replicas in that Shepherd.

\subsubsection{Replication}
\label{Replication}
Shepherds periodically ask the Librarian if the file of the replica stored on
its storage node have enough replicas. If the file does not have
enough replicas, the Shepherd informs the Bartender and the Bartender initiates
a put request that returns a Transfer URL to the Shepherd. The
Shepherd finally uploads the new replica. The Shepherds which gets the new replica notifies the
Librarian that the replica is alive. The Librarian then add this to
the corresponding file.

@@

Due to the unpredictable nature of the Grid environment, it is essential to
have some degree of  automatic recovery system in case of unexpected
failures. Fault tolerant behavior is required both at the level of
metadata and on the level of physical storage. While the work on fault-tolerant metadata is still in
progress, the following two
scenarios will explain the currently available recovery mechanism for
physical storage: 
\begin{itemize}
\item In the case of a file having invalid checksum, the Shepherd
immediately informs the Librarian and the Librarian changes the state of the given replica to
`\textbf{invalid}'. To recover its replica, the Shepherd contacts a
Bartender and asks for another replica of the file. The Bartender chooses a valid
replica, initiate file transfer from a Shepherd having the replica,
and returns the TURL to the Shepherd with the invalid replica. When the
Shepherd has recieved the replica and compared the checksum, it notifies the
Librarian that the replica is alive again.  
\item In the case of a Shepherd becoming offline, a Librarian will
  as mentioned notice the lack of heartbeats and invalidate all the
  replicas, initiating new replication for all the files stored in
  this storage node. However, if the Shepherd
  again comes online, there will evidently be more replicas than
  needed. The first Shepherd to notice this will set its replica's
  state to `\textbf{thirdwheel}', i.e., obsolete. At the next occasion,
  the Shepherd will remove the replica, if and only if it has the only
  `\textbf{thirdwheel}' replica of this file. If there are more
  replicas with this state, all replicas will be set back to
  `\textbf{alive}' and the process is repeated. This scenario will be discussed further in Section
  \ref{Testing and Discussion}.
\end{itemize}
\subsubsection{Client tools }
\label{Client extension using FUSE module}

Being the only part a user will (and should) see from a storage
system, the client tools is an important part of the ARC
Storage. Currently ARC supports two ways of accessing the
storage solution:
\begin{itemize}
\item The \textbf{Command-line Interface} (CLI) provides
  access to the storage through the methods \texttt{stat}, \texttt{makeCollection},
  \texttt{unmakeCollection}, \texttt{putFile}, \texttt{getFile}, \texttt{delFile},
  \texttt{list} and \texttt{move}. Methods for modifying
  access and ownership will be available in the near future. The CLI
  does assume a relatively high level of computer experience
  from the user. However, the CLI, being a stand-alone tool, can be
  used to access ARC Storage from any computer\footnote{Also including
  Microsoft Windows PCs.} that has network access and
  a Python installation.
\item The \textbf{FUSE module} provides a high-level access to the
  ARC Storage. Filesystem in Userspace (FUSE) provides a simple
  library and a kernel-userspace interface. Using FUSE and the ARC
  Storage Python interface, the FUSE module allows users to mount the
  ARC Storage namespace into the local namespace. This way, the user
  can use his/her favorite file manager to access his/her files and
  collections. The FUSE module provides most of the features provided
  by the CLI, with the exception of modifying some non-posix metadata.
\end{itemize}
Its worth mentioning that the client tools access the storage system
through the Bartender only. Currently upload and download is realized
through HTTP(S), but there are plans to add support for other
protocols, such as SRM and GridFTP.

@@

Gateways are used to communicate with the external storage
managers. While designing this service, care was taken to: 
\begin{itemize}
\item Retain the transparency of the global namespace while using the external storage systems.
\item  Develop a protocol oriented  service i.e., all the external
  storage managers which support a certain protocol should be handled
  using the corresponding gateway service.  
\end{itemize}
This approach provides the flexibility of  avoiding multiple Gateway
services for different storage managers. Currently the available
Gateway service is based on the gridftp protocol.

When the user requests comes, the Librarian provides the metadata related
with the request to the Bartender. Requests can be related to
files, collections or external mount-points. In case of creating
external mount-points, the Bartender contacts the Librarian to store
the mount-point for later use. In the case where a request is related
to the downloading of files from the external store, the Gateway service first checks the status of the file
and then send the Transfer URL (TURL) to the client via the
Bartender. Using this TURL client can directly get the file from
external store. %% For the experiments we use dCache with gridftp
%% protocol as an external storage system.

@@

The services were distributed as shown in figure \ref{fig:map}:
\begin{itemize}
\item An A-Hash runs at UPPMAX.
\item A Bartender runs at USIT.
\item A Librarian runs on UPPMAX.
\item In total five Shepherds were used for the tests: Three at USIT, having
  100GB storage space each, one at UPPMAX, with 20GB, and one at NIIF, providing 16GB of storage space.
 \end{itemize}

@@

Figure \ref{fig:FileDistribution} illustrate two test results,
corresponding to Test 2 and Test 3: The orange bars show the distribution after uploading 100 small files
simultaneously, without replication. This gives an indication of how
the system balances the load of many clients uploading at the same
time. The green bars show distribution with the same kind of
uploading, but with the clients requesting three replicas for each file. The
difference here is that the system simultaneously has to handle both
replication and the clients requesting uploads. When the Bartender chooses
a location for which to put a replica, it generates a list of
Shepherds that (a) do not have a replica of the file, and (b) is not
already in the process of uploading the replica. Next, it draws a
Shepherd from the list at random, using a uniform random number
generator. When uploading without asking for replicas we would
therefore expect a relatively flat distribution of the files, as can
be seen in figure \ref{fig:FileDistribution}.

>>>>

\begin{document}
\pagestyle{headings} 
\title{Service Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$  \\ $^{*}$Scientific
  Computing Division, IT Department, Uppsala University, Sweden
  \\ $^{\dagger}$Center for Information Technology, University of Oslo, Norway
  \\ $^{\ddagger}$Department of Physics,  University of Oslo, Norway
  \\  $^{\mathsection}$Department of Physics and Astronomy, Uppsala University,
  Sweden \\ $^{\mathparagraph}$National Information Infrastructure Development
  Institute, Hungary\\ \\  j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

\maketitle             
%% \begin{abstract}
%% The KnowARC project will deliver the next generation Advanced Resource
%% Connector Fall  2009. A major component of this
%% service oriented Grid middleware will be its storage system, the
%% ARC Storage, of which a
%% prototype is due Spring 2009. We will in this paper present an
%% overview of the storage system, its design and architecture. We will
%% also present some first, proof-of-concept test results, deploying the
%% the storage with a geographically distribution among three different countries.
%% ~\\~\\~
%% \end{abstract}
\begin{abstract}
There is an ever increasing need to utilize geographically distributed
hardware resources, both in terms of CPU and in terms of storage. The
service oriented architecture provides a natural framework for
managing these resources. The next generation Advanced Resource
Connector (ARC) is a service oriented Grid solution that will provide
the middleware to represent distributed resources in one simple
framework. In this paper, we will present an overview of the ARC
storage system, itself a set of services providing a self-healing,
flexible Grid storage solution. We will also present some first
proof-of-concept test results, with a deployment of the storage system
distributed across three different countries. 
~\\~\\~
\end{abstract}
%-----------------------------------------------------------
\section{Introduction}
\label{Introduction}

The challenge of building a reliable, self-healing, fault-tolerant,
consistent data management system in a web scale is an interesting
task. Making the system work in a heterogeneous, distributed
environment like the Grid is even more interesting. An increasing
number of applications demand not only increased CPU power, but also
vast amounts of storage space. The required storage space is not only
restricted to the duration in which the application runs; the data
should often be available for years afterwards, in certain cases even
for decades. Nowadays, we can easily find single Grid jobs which produce
gigabytes or even terabytes of data, ramping up the requirements of
storage systems to the petabyte scale and beyond. Hard drive capacity is
still increasing and storage raids can handle failing hard drives through 
data replication. However, even a designated storage site cannot guarantee 
24/7 availability or long term security against catastrophic losses of data. 
The need for a distributed, self-healing storage system is evident. To
make the storage system useable to Grid users, the system must provide
reliable and secure file transfer protocols, a cataloging system and
secure storage of the data. Several projects and designs have emerged
to address such challenges \cite{Hoschek00datamanagement,DengWang}.

In the advent of the next generation Advanced Resource Connector (ARC)
Grid middleware \cite{NorduGridsite} (due during Fall 2009), we present 
the ARC storage system \cite{ARCStoragedesigndoc}. This distributed
storage system is designed to provide an easy
to use, flexible and scalable system that can offer native storage and
at the same time provide access to third-party solutions like dCache \cite{dCache,DSSWithdCache}  by
using the same, uniform interface. Being part of ARC, the
storage system is based on a service
oriented architecture, in which each major component of the
system runs as a separate service within the ARC Hosting Environment
Daemon (HED) \cite{HEDdesigndoc}. The HED service container gives the
capability of flexible replacement of the components as well as the
possibility to introduce modifications in the future. %% The HED service container gives
%% the capability of flexible replacement of the components as well as
%% the future modification/enhancement of the system.

This paper is organized as follows. After describing
related work in Section \ref{Related Work}, we give a bird's eye
view of the next generation Advanced Resource Connector (ARC) in
Section \ref{The Advance Resource Connector}. An overview of the
ARC Storage is given in Section \ref{The ARC Storage System}, while
the architecture of the storage system is elaborated in Section
\ref{Architecture of the ARC Storage}. In Section \ref{Testing and
  Discussion} we give some early, proof-of-concept results, before 
concluding in Section \ref{Conclusion and Future Work}.

@@

\textbf{dCache:} dCache is a storage system which combines heterogeneous storage

@@

a joint effort between DESY \cite{dCachesite}, Fermi National Accelerator Laboratory \cite{FermiLab}, the  Nordic Data

@@

it is relatively difficult to deploy and integrate with new applications. 
The ARC Storage, being a light-weight and flexible

@@

\textbf{BigTable:} Bigtable is a distributed storage system managing structured data on

@@

Google in projects like web indexing, Google Earth and Google

@@

system. It allows  uniform access to heterogeneous

@@

Originally developed for use with the physics analysis tool root \cite{root},

@@

high bandwidth environments. In contrast, ARC Storage is optimized
for high latency more suitable for the Grid environment.

@@

The next generation Advanced Resource Connector (ARC) Grid middleware 
\cite{NorduGridsite} \cite{KnowARCsite} consists of a set of pluggable 
components. These components are the
fundamental building blocks of the ARC services and clients. ARC
services run inside a container called the Hosting Environment
Daemon (HED) and there are four kinds of pluggable components with well
defined tasks: Data Management Components are used to transfer the
data using various protocols, Message Chain Components are
responsible for the communication within clients and services as well as
between the clients and the services, ARC Client Components are
plug-ins used by the clients to connect to different Grid flavors, and
Policy Decision Components are responsible for the security model
within the system.

To deliver the non-trivial quality of services required by the Grid,
there are a number of services running inside the HED. For example, Grid
job execution and management is handled by the A-REX service 
\cite{AREXdesigndoc}, policy
decisions are taken by the Charon service, the ISIS service is
responsible for the information indexing, batch job submission is
handled by the Sched service, etc. In the later sections our discussion
will focus on the architecture and the design of another major set
of services, i.e., the ARC storage.

@@

The ARC Storage consists of a set of SOAP based services residing
within HED. Together, the services provide a self-healing, reliable, robust,
scalable, resilient and consistent data storage system. Data is
managed in a hierarchical global namespace with files and
subcollections grouped into collections\footnote{A concept very
  similar to files and directories in most common file systems.}.  A
dedicated root collection serves as a reference point for accessing the namespace. The hierarchy can then be
referenced using Logical Names. The global namespace is accessed
in the same manner as in local filesystems.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.8\columnwidth]{arc1-storage-services-with-gateway.pdf}
  \end{center}
  \caption{Schematic of the ARC Storage architecture. The figure shows
    the main services of ARC Storage and the communication channels. The B stands for Bartender, the S Shepherd, the L Librarian, the G Gateway and the A-H stands for A-Hash. The straight lines denotes the communication between services.}
  \label{fig:arc1_storage_service}
\end{figure}

Being based on a service oriented architecture, the ARC
Storage consists of a set of services as shown in Fig.
\ref{fig:arc1_storage_service}. The services are as follows: The Bartender (B)
provides the high-level interface to the user; the Librarian (L)
handles the entire storage namespace, using the A-Hash (A-H) as a metadatabase; the
Gateway (G) provides access to third-party storage systems; and
the Shepherd (S) is the frontend for the physical storage node. 
See Section \ref{Architecture of the ARC Storage} for a detailed
discussion of the different services.   
The services communicate with each other through the
Message Chain Components in HED. The communication channels are
depicted by straight lines in Fig. \ref{fig:arc1_storage_service}.

The system supports file transfer through several transfer protocols,
with client side tools that hide various technical
details such as protocol specification, port numbers
and so on. To provide fault-tolerance, the system implements
automatic file replication, where
the replicas of a file\footnote{In this paper, a file
denotes an entry in the global namespace, while replica denotes the
physical file stored on a storage node.} are always stored on separate
storage nodes\footnote{A
  storage node is a server with a Shepherd, a Shepherd backend
and some storage service.}. To ensure a resilient, self-healing system, the Shepherd
regularly sends heartbeats to one of the Librarians. If the Shepherd fails
to send a heartbeat, one of the Librarians will automatically initiate
re-replication of the replicas between the other Shepherds.

In the default implementation, the services will provide a
full-featured and consistent data storage system using files as an
atomic unit. The scope of the ARC storage system in the foreseen
future is restricted to provide support for file-based data services.
%% there are currently no plans for intrinsic support for database
%% access.

@@

External grid middleware components can
access the ARC storage system by using ARC data service
interfaces directly. ARC will also provide interface
components that communicate via standard protocols like SRM, which
will provide a single access point to the whole system.

@@

In a service oriented architecture
the role of each service is well defined. Available
objects\footnote{Files, collections, mount points, etc.} in the
system are identified by unique global IDs. These IDs are
categorized according to the object type:

\begin{itemize} 
\item Each  file and collection have a unique ID called GUID.
\item Services are uniquely identified by a serviceID. 
\item The Shepherds in the system identify their files by a referenceID.
\end{itemize}

Each object in the ARC Storage has a globally
unique ID. A collection contains files and other collections,
and each of these entries has a name unique within the collection very
much like entries in a standard directory on a local filesystem. Besides
files and collections the ARC Storage has a third type of entry
called mount-points, which are references to the third-party storages
within the global namespace.

Replicas in a distributed storage system can have different states;
they can be broken, deleted, partially uploaded, etc. In the
ARC Storage, all replicas have assigned a state, some of which are
`\textbf{alive}' (if the replica passed the checksum test, and the
Shepherd reports that the storage node is healthy), `\textbf{invalid}'
(if the replica has wrong checksum, or the Shepherd claims it has no
such replica), `\textbf{creating}' (if the replica is in
the state of uploading).

In the following the details of the
architecture are presented in three parts: First we will discuss the
details of the core components, second we will discuss some of the
important features provided by the system, and third we will discuss the security model.

\subsection{Core Components} 
\subsubsection{Bartender}
\label{Bartender}
The Bartender service provides a high-level interface for the storage
system. Clients connect to the Bartender to create and remove
files, collections and mount-points using their Logical Names. The
Bartender communicates 
with the Librarian and Shepherd services to execute the clients'
requests. However, the actual file data does not go through the
Bartender, instead file
transfers are directly performed between the storage nodes and the
clients. There could be any number of independent Bartender services
running in the system, providing high-availability and
load-balancing. 
\subsubsection{Librarian}
\label{librarian}
The Librarian manages the hierarchy and metadata of
files, collections and mount points, as well as the health information of the Shepherd
services. In addition, the Librarian handles
the information about registered Shepherd services. The Librarian receives
heartbeat messages from the Shepherds and change replica states automatically
if needed. The Librarian uses the A-Hash for consistently storing all metadata. This makes the
Librarian a stateless service, thus enabling the system to have any
number of independent
Librarian services, again providing high-availability in the
system.

@@

The Shepherd services run as front-ends on storage nodes. A Shepherd
service reports to a Librarian about the 
node's health state in terms of replicas. While the Bartender initiates
file transfers, the actual transfers go directly between the Shepherd
and the clients.

@@

A-Hash is a hash table for consistently storing data in property-value
pairs. All metadata about files, collections, mount point, Shepherd's
health status, and so forth, is stored in the A-Hash. As the A-Hash is
the service storing the entire state of the storage system, it is
absolutely crucial for the ARC Storage that the A-Hash is
consistent. The distribution and replication of this service is
therefore both necessary and challenging.

@@

\subsection{Features}
\subsubsection{Heartbeat monitoring}
\label{Heartbeat and Replication}
In the proposed architecture, each Shepherd periodically sends
heartbeats to a Librarian with information about replicas
whose state changed since the last heartbeat, and the corresponding
GUIDs. These heartbeats are then stored in the A-Hash, making them
visible to all the Librarians in the system. If any of the Librarians
notices that a Shepherd is late with its heartbeat, it will invalidate
all the replicas in that Shepherd.

\subsubsection{Replication}
\label{Replication}
Shepherds periodically ask the Librarian if the file of the replica stored on
its storage node has enough replicas. If the file does not have
enough replicas, the Shepherd informs the Bartender and the Bartender initiates
a put request that returns a Transfer URL to the Shepherd. The
Shepherd finally uploads the new replica. The Shepherds which gets the new replica notifies the
Librarian that the replica is alive. The Librarian then add this to
the corresponding file.

@@

Due to the unpredictable nature of the Grid environment, it is essential to
have some degree of  automatic recovery system in case of unexpected
failures. Fault tolerant behavior is required both at the level of
metadata and on the level of physical storage. While the work on fault-tolerant metadata is still in
progress, the following two
scenarios will explain the currently available recovery mechanism for
physical storage: 
\begin{itemize}
\item In the case of a file having invalid checksum, the Shepherd
immediately informs the Librarian and the Librarian changes the state of the given replica to
`\textbf{invalid}'. To recover its replica, the Shepherd contacts a
Bartender and asks for another replica of the file. The Bartender chooses a valid
replica, initiates a file transfer from a Shepherd having the replica,
and returns the TURL to the Shepherd with the invalid replica. When the
Shepherd has recieved the replica and compared the checksum, it notifies the
Librarian that the replica is alive again.  
\item In the case of a Shepherd going offline, a Librarian will,
  as mentioned earlier, notice the lack of heartbeats and invalidate all the
  replicas, initiating new replication for all the files stored in
  this storage node. However, if the Shepherd
  again comes online, there will evidently be more replicas than
  needed. The first Shepherd to notice this will set its replica's
  state to `\textbf{thirdwheel}', i.e., obsolete. At the next occasion,
  the Shepherd will remove the replica, if and only if it has the only
  `\textbf{thirdwheel}' replica of this file. If there are more
  replicas with this state, all replicas will be set back to
  `\textbf{alive}' and the process is repeated. This scenario will be discussed further in Section
  \ref{Testing and Discussion}.
\end{itemize}
\subsubsection{Client tools }
\label{Client extension using FUSE module}

Being the only part a user will (and should) see from a storage
system, the client tools are an important part of the ARC
Storage. Currently ARC supports two ways of accessing the
storage solution:
\begin{itemize}
\item The \textbf{Command-line Interface} (CLI) provides
  access to the storage through the methods \texttt{stat}, \texttt{makeCollection},
  \texttt{unmakeCollection}, \texttt{putFile}, \texttt{getFile}, \texttt{delFile},
  \texttt{list} and \texttt{move}. Methods for modifying
  access and ownership will be available in the near future. The CLI
  assumes a relatively high level of computer competence
  from the user. However, the CLI, being a stand-alone tool, can be
  used to access ARC Storage from any computer (also including
  Microsoft Windows PCs) that has network access and
  a Python installation.
\item The \textbf{FUSE module} provides a high-level access to the
  ARC Storage. Filesystem in Userspace (FUSE) \cite{FUSE} provides a simple
  library and a kernel-userspace interface. Using FUSE and the ARC
  Storage Python interface, the FUSE module allows users to mount the
  ARC Storage namespace into the local namespace. This way, the user
  can use her/his favorite file manager to access her/his files and
  collections. The FUSE module provides most of the features provided
  by the CLI, with the exception of modifying some non-posix metadata.
\end{itemize}
Its worth mentioning that the client tools access the storage system
through the Bartender only. Currently upload and download is realized
through HTTP(S), but there are plans to add support for other
protocols, such as SRM and GridFTP.

@@

Gateways are used to communicate with the external storage
managers. While designing this service, care was taken to: 
\begin{itemize}
\item Retain the transparency of the global namespace while using the external storage systems.
\item  Develop a protocol oriented  service i.e., all the external
  storage managers which support a certain protocol should be handled
  using the corresponding gateway service.  
\end{itemize}
This approach provides flexibility while avoiding multiple Gateway
services for different storage managers. Currently, the available
Gateway service is based on the gridftp protocol.

When the user request is made, the Librarian provides the metadata related
to the request to the Bartender. Requests can be related to
files, collections or external mount-points. In case of creating
external mount-points, the Bartender contacts the Librarian to store
the mount-point for later use. In the case where a request is related
to the downloading of files from the external store, the Gateway service first checks the status of the file
and then sends the Transfer URL (TURL) to the client via the
Bartender. Using this TURL, the client can directly get the file from
external store. %% For the experiments we use dCache with gridftp
%% protocol as an external storage system.

@@

The services were distributed as shown in Fig. \ref{fig:map}:
\begin{itemize}
\item An A-Hash runs at UPPMAX.
\item A Bartender runs at USIT.
\item A Librarian runs on UPPMAX.
\item In total five Shepherds were used for the tests: Three at USIT, having
  100GB storage space each, one at UPPMAX, with 20GB, and one at NIIF, providing 16GB of storage space.
 \end{itemize}

@@

Fig. \ref{fig:FileDistribution} illustrate two test results,
corresponding to Test 2 and Test 3: The orange bars show the distribution after uploading 100 small files
simultaneously, without replication. This gives an indication of how
the system balances the load of many clients uploading at the same
time. The green bars show distribution with the same kind of
uploading, but with the clients requesting three replicas for each file. The
difference here is that the system simultaneously has to handle both
replication and the clients requesting uploads. When the Bartender chooses
a location for which to put a replica, it generates a list of
Shepherds that (a) do not have a replica of the file, and (b) is not
already in the process of uploading the replica. Next, it draws a
Shepherd from the list at random, using a uniform random number
generator. When uploading without asking for replicas we would
therefore expect a relatively flat distribution of the files, as can
be seen in Fig. \ref{fig:FileDistribution}.
